---
title: "Rmarkdown Practice with Penguin Dataset"
author: "Retno K. Ningrum"
date: "2024-10-01"
output: 
  html_document:
    toc: TRUE  #if you want to create a table of content
    toc_float: TRUE #the style how you want to show the table of content
    theme: darkly #using the theme, you can explore more in the global options 
  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../output/")
```

# Intro to Penguins
today we are going to work with penguin dataset using the rmarkdown to learn **rmarkdown**

### load libraries
```{r}
library(palmerpenguins)
library(tidyverse)
library(here)
library(ggplot2)
```
### inspect the data
```{r}
glimpse(penguins)

```

### make some plot 
```{r}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```

##Change size
#you can change size in the begining of the chunk "70%" means makes it 70% of the width
```{r, out.width= "70%"}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```

## change size with more specific value
```{r, fig.width=3, fig.height=4}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```

## change size using ratio AND center the alignment
```{r, fig.asp=1.2, fig.align='center'}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```

##NOTES
# you need to have 1 figure in each R chunk
# you need to rename your code chunk
# now lets re-do! and rename with "penguinplot"
#Rmarkdown cannot use "here" package, so in order to save the output, you have to set it up in the first chunck to re-direct the output. I write the begining of the cunk as   fig.path = "../output/"
#meaning that "go back to the outer folder, then get inside to the output folder
#once the chunck is set, then all the chunk can be rename into different figures name after r code in the first chunk
# now lets re-do! and rename with "pengunplot"

```{r penguinplot, fig.asp=1.2, fig.align='center'}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```

#Adding caption to your figures
```{r penguin_caption , fig.cap= "fugure xx. a scaterplot of penguin" }
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```

#save into pdf
```{r penguin_caption2 , fig.cap= "fugure xx. a scaterplot of penguin", dev="pdf" }
penguins %>%
  ggplot(aes(x = bill_length_mm, y = bill_depth_mm, color= species)) +
  geom_point()+
  geom_smooth(method = "lm") +
  labs(x = "Bill length (mm)", y = "Bill depth (mm)") +
  theme_bw()
```



| Time | Session | Topic |
|:-----|:--------|:------|
|_left_|_center_|_right_|
|01:10 |1       |Anatomy|
|01:00 |2       |*Break*|
|01:12 |2       |Tables |

```{r}
## to make a summary table directly, use kabletable
library(kableExtra)
library(tidyverse)

penguins %>% 
  group_by(species) %>%
  Summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() # make it a kable table, very basic

```


## Kable Table Package
```{r}
penguins %>% 
  group_by(species) %>%
  Summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() +                        # make it a kable table
  kable_classic() %>%            #adding the theme in the kable
  row_spec(2, bold =  T, color = "white", background = "red") %>%  #Highlight to row 2, where background color is red, and the text color is white. 
  kable_styling(full_width = FALSE) #If you want the table not too wide
```

