---
title: 'Tidy Tuesday : ISO'
author: "Retno K. Ningrum"
date: "2024-11-11"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,                  #display the code
                      message = FALSE,              #do not show the message
                      warning = FALSE,              #do not show the warning
                      fig.path = "../output/")      #set the output for image
```

## Load all Libraries
```{r}
library(countries)
library(dplyr)
library(ggplot2)

```

## Read Data and Check Data
```{r}

#Read the data
countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/countries.csv')
country_subdivisions <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/country_subdivisions.csv')
former_countries <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-11-12/former_countries.csv')


#Check the data
glimpse(countries)
glimpse(country_subdivisions)
glimpse(former_countries)

```

## Data Manipulation
After check the data, I decide to just use the Countries and Country_subdivision data to visualize the total subdivision of each country in the world.
```{r}
# Join countries and country_subdivisions
all_countries <- full_join(countries,              #join countries and country_subdivision dataset
                           country_subdivisions, 
                           by = "alpha_2") %>%     #using alpha_2 as the reference to join the dataset
  group_by(name.x, alpha_3) %>%                    #group by name.x (name country) and alpha_3
  summarise(total = n())                           #create new column to store the total row found

```
## Create the Map
I will try new package {countries} and use the quick_map() function to generate the map of total subdivision each country in the world
```{r Map_Subdivision, fig.width= 12, fig.height=5, fig.align= 'center'}
# Create the map
quick_map(all_countries,       #create map, using all_countries data
          plot_col = "total",  #use the value in "total" column 
          theme = 9) +         #choose viridis color to fill the map, represent total value
  coord_map("mercator",          #choose the mercator type projection  
            xlim = c(-180, 180), #set the xlim
            ylim = c(-60, 80)) + #set the ylim
  labs(title = "Country Subdivision in the World",   #add title
       caption = "Source: Tidy Tuesday ISO Dataset", #add caption
       fill = "Total                                       
Subdivision") +
  theme(plot.title = element_text(size = 14,                #set the font size of the title
                                  hjust = 0.5,                #set the position of title
                                  face = "bold",            #set title into bold
                                  margin = margin(b = 20)), #set margin between title and map
        plot.caption = element_text(size = 10)) +   #add caption, set size
   theme(legend.position = "right",                #Move legend to the right
         legend.title = element_text(size = 10),    #set the size of legend title
         legend.text = element_text(size = 10))     #set the size of legend text

```

## New things I learned:  
1. quick_map() function. This is the first time I try quick_map to create a map. The data needed to create map is only the country's name and the value on each country.  
2. Margin in theme(). This is the first time I use margin, which allow me to set the distance between plot title to the plot/map. 

