---
title: "Homework Week 7"
author: "Retno K. Ningrum"
date: "2024-10-09"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,                  #display the code
                      message = FALSE,              #do not show the message
                      warning = FALSE,              #do not show the warning
                      fig.path = "../output/")      #set the output for image
```
# Creating Map with R

This assignment use [Plastic Pollution Dataset](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv) with specific countries in Southeast Asia (*Indonesia, Singapore, Malaysia, Philippines, Thailand, Vietnam, Cambodia, Laos, China,Timor-Leste, Myanmar*). 

## Overall Steps Summary 
1. Filter data plastic from Southeast Asia (SEA) country and then sum all the total plastic in each country
2. Filter world data using country mentioned in plastic data in SEA
3. Join the plastic data in SEA with the world data
4. Create text for each country label manually, following the coordinate of each country
5. create the plot!

### Load Libraries
load all necessary libraries
```{r}

library(tidyverse) 
library(here)
library(maps)
library(mapdata)
library(mapproj)


```

### Load and Inspect the Data
```{r}
#Read the plastic data
plastics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-01-26/plastics.csv')

#Read the world data (for mapping)
world <- map_data("world")

#Check the first 10 data
head(plastics)
head(world)

#check overall data
glimpse(plastics)
glimpse(world)

unique(plastics$country)  #to check the list of country covered in the plastics data
unique(world$region)      #to check the list of country covered in the world data

```

After checking the dataset, I decide to focus on creating map of plastic distribution in Southeast Asia

### Data Cleaning
```{r}
#create the SouthEast Asia (SEA) dataset for map background
SEA <- world %>%                                                    #use world data
  filter(region %in% c("Indonesia", "Singapore", "Malaysia",        #filter data in region column
                       "Philippines", "Thailand", "Vietnam", 
                       "Cambodia", "Laos", "China", "Timor-Leste", 
                       "Myanmar")) %>%
  select(long, lat, group, region)                          #select only long, lat, group, region column


#create the plastic data in SEA
plastic_SEA <- plastics %>%                          #use plastic data
  filter(country %in% c("Indonesia", "Singapore",    #filter country in SEA
                        "Malaysia", "Philippines", 
                        "Thailand", "Vietnam")) %>%          
  select("region" = country,  #Select the country column, but rename it to "region" -> for join purposes
         grand_total) %>%     #select the grand_total value    
  group_by(region)  %>%       #group by region
  summarise(total_plastics = sum(grand_total))  #summarize grand_total in each country/region

#Join the plastic in SEA with SEA
SEA_plastic <- left_join(SEA, plastic_SEA)

#Create label for the map
country_labels <- data.frame(
  region = c("Indonesia", "Malaysia", "Philippines",  #show the label for these countries
             "Singapore", "Thailand", "Vietnam"),
  long = c(113.9213, 101.9758, 122.5583, 103.8198, 100.9925, 105.8562), # set longitudes of each country label
  lat = c(-0.7893, 4.2105, 12.8797, 1.3521, 15.8700, 14.0583)        # set latitude if each country label
)

```

### Plot the Map!
```{r plastic_SEA}

# Plot the map
ggplot(data = SEA_plastic) +                          #use data world_indo as the plot background
  geom_polygon(aes(x = long, y = lat,                 #create polygon using x.y coordinate
                   group = group,                     #group by group
                   fill = total_plastics),            #fill data using total_plastics
               color = "white") +                     #country border is white
  geom_text(data = country_labels,                    #use data from country_labels
            aes(x = long, y = lat,                    #set text by long and lat axis
                label = region),                      #label text by region
            size = 3.5,  color = "black") +           #set the size by 3.5, text color is black
  coord_map(projection = "mercator", #Use the mercator type as projected map
            xlim = c(95, 139),       #set the x coordinate from 95 to 139 (this is indonesia coordinate) 
            ylim = c(-10, 21)) +     #set the y coordinate from -10 to 12 (this is indonesia coordinate)
  labs(title = "Plastic Distribution in Southeast Asia",  #Add title
       subtitle = "Indonesia, Singapore, Malaysia, Philippines, Thailand, Vietnam",   #Add subtitle
       x = "Longitude",                                   #Add X-axis label
       y = "Latitude",                                    #Add Y-axis label
       fill = "Total Plastics") +                         #Add legend title
  theme_minimal() +                                       #Apply minimal theme for clean appearance
  theme(panel.grid.major = element_line(color = "gray",       #Add gridline, color grey
                                        linetype = "dotted"), #set type of gridline: dotted gridline
        plot.title = element_text(hjust = 0.5,     #place the title text in the center
                                  size = 16,       #Set font size = 16
                                  face = "bold"),
        plot.subtitle = element_text(hjust = 0.5,  #place the subtitle text in the center
                                  size = 12 ))     #Set font size = 12
                                

```

